### 2022-06-13 doc/src/IMSIC.tex
Guest interrupt files do not support value 0x40000000 for eidelivery

For machine level, the relevant CSRs are:\nopagebreak
\begin{displayLinesTable}[l@{\qquad}l@{\qquad}l]
\z{miselect} & \z{mseteipnum} & \z{mtopei} \\
\z{mireg}    & \z{mclreipnum} \\
             & \z{mseteienum} \\
             & \z{mclreienum} \\
\end{displayLinesTable}

When supervisor mode is implemented, the set of supervisor-level CSRs
matches those of machine level:
\begin{displayLinesTable}[l@{\qquad}l@{\qquad}l]
\z{siselect} & \z{sseteipnum} & \z{stopei} \\
\z{sireg}    & \z{sclreipnum} \\
             & \z{sseteienum} \\
             & \z{sclreienum} \\
\end{displayLinesTable}

And when the hypervisor extension is implemented, there is a
corresponding set of VS CSRs:
\begin{displayLinesTable}[l@{\qquad}l@{\qquad}l]
\z{vsiselect} & \z{vsseteipnum} & \z{vstopei} \\
\z{vsireg}    & \z{vsclreipnum} \\
              & \z{vsseteienum} \\
              & \z{vsclreienum} \\
\end{displayLinesTable}

@@

The indirectly accessed interrupt-file registers, and the other CSRs
that directly interact with external interrupts (\z{mseteipnum},
\z{mclreipnum}, etc.), are all documented in more detail in the next
two sections.

@@

%-----------------------------------------------------------------------
\section{CSRs directly affecting an interrupt file}

Several new machine-level CSRs interact directly with an IMSIC's
machine-level interrupt file:  \z{mseteipnum}, \z{mclreipnum},
\z{mseteienum}, \z{mclreienum}, and \z{mtopei}.
If supervisor mode is implemented, there are several matching
supervisor CSRs that likewise interact with the supervisor-level
interrupt file:  \z{sseteipnum}, \z{sclreipnum}, \z{sseteienum},
\z{sclreienum}, and \z{stopei}.
And if the hypervisor extension is implemented and field VGEIN of
\z{hstatus} is the number of an implemented guest interrupt file,
several VS CSRs interact with the chosen guest interrupt file:
\z{vsseteipnum}, \z{vsclreipnum}, \z{vsseteienum}, \z{vsclreienum},
and \z{vstopei}.
These CSRs are documented by function below:

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{%
Set/clear external interrupt-pending bit by number
(\zSafe{*seteipnum} and \zSafe{*clreipnum})%
}

If $i$ is an implemented interrupt identity number in the respective
interrupt file, writing value $i$ to a \emph{\texttt{*seteipnum} CSR}
(\z{mseteipnum}, \z{sseteipnum}, or \z{vsseteipnum}) causes the pending
bit for interrupt~$i$ to be set to one in the interrupt file, while
writing $i$ to a \emph{\texttt{*clreipnum} CSR} (\z{mclreipnum},
\z{sclreipnum}, or \z{vsclreipnum}) causes the pending bit for
interrupt~$i$ to be cleared in the interrupt file.
A write to a \z{*seteipnum} or \z{*clreipnum} CSR is ignored if the
value written is not an implemented interrupt identity number for the
respective interrupt file.

A read of a \z{*seteipnum} or \z{*clreipnum} CSR always returns zero.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{%
Set/clear external interrupt-enable bit by number
(\zSafe{*seteienum} and \zSafe{*clreienum})%
}

If $i$ is an implemented interrupt identity number in the respective
interrupt file, writing value $i$ to a \emph{\texttt{*seteienum} CSR}
(\z{mseteienum}, \z{sseteienum}, or \z{vsseteienum}) causes the enable
bit for interrupt~$i$ to be set to one in the interrupt file, while
writing $i$ to a \emph{\texttt{*clreienum} CSR} (\z{mclreienum},
\z{sclreienum}, or \z{vsclreienum}) causes the enable bit for
interrupt~$i$ to be cleared in the interrupt file.
A write to a \z{*seteienum} or \z{*clreienum} CSR is ignored if the
value written is not an implemented interrupt identity number for the
respective interrupt file.

A read of a \z{*seteienum} or \z{*clreienum} CSR always returns zero.

%- - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
\subsection{Top external interrupt (\zSafe{*topei})}

@@

If it is necessary first to read a \z{*topei} CSR and then subsequently
claim the interrupt as a separate step, the claim can be safely done by
writing the interrupt identity to the corresponding \z{*clreipnum} CSR,
instead of writing to \z{*topei}.
\end{commentary}

\begin{commentary}
Because the effect of a combined read-and-write of a \z{*topei} CSR
can be accomplished equally by a read of \z{*topei} followed by a write
of the interrupt identity to the corresponding \z{*clreipnum}, the
\z{*topei} CSRs could have been made read-only without the special
behavior defined for writes.
However, condensing these operations into a single CSR access instead of
two helps to reduce interrupt latency when critical.
\end{commentary}

>>>>

For machine level, the relevant CSRs are:\nopagebreak
\begin{displayLinesTable}
\z{miselect} \qquad \z{mireg} \qquad \z{mtopei} \\
\end{displayLinesTable}

When supervisor mode is implemented, the set of supervisor-level CSRs
matches those of machine level:
\begin{displayLinesTable}
\z{siselect} \qquad \z{sireg} \qquad \z{stopei} \\
\end{displayLinesTable}

And when the hypervisor extension is implemented, there is a
corresponding set of VS CSRs:
\begin{displayLinesTable}
\z{vsiselect} \qquad \z{vsireg} \qquad \z{vstopei} \\
\end{displayLinesTable}

@@

The indirectly accessed interrupt-file registers
and CSRs \z{mtopei}, \z{stopei}, and \z{vstopei}
are all documented in more detail in the next
two sections.

@@

%-----------------------------------------------------------------------
\section{%
Top external interrupt CSRs (\zSafe{mtopei}, \zSafe{stopei}, \zSafe{vstopei})%
}

CSR \z{mtopei} interacts directly with
an IMSIC's machine-level interrupt file.
If supervisor mode is implemented, CSR \z{stopei} interacts
directly with the supervisor-level interrupt file.
And if the hypervisor extension is implemented and field VGEIN of
\z{hstatus} is the number of an implemented guest interrupt file,
\z{vstopei} interacts with the chosen guest interrupt file.

@@

If it is necessary first to read a \z{*topei} CSR and then subsequently
claim the interrupt as a separate step, the claim can be safely done by
clearing the pending bit in the \z{eip} array
via \z{*siselect} and \z{*sireg},
instead of writing to \z{*topei}.
\end{commentary}
